Data preparation

### 2025-07-29 procesamiento/02-prod_prep-country.Rmd
ajustes data micro

```{r merge}
macro_data <- 
  left_join(wiid_2017,solt_2017) %>%
  left_join(pen_2017) %>% 
  left_join(otero) %>%
  left_join(ocde) %>% 
  left_join(wid_df1) %>% 
  left_join(ilo_df)

save(macro_data,file = here::here("input/data/proc/study1_macro_data.RData"))

@@

```{r}
summary(df1$incomepcap)

df_class3_ineq <- 
df2 %>%   
  group_by(country2,dclass3res_V) %>%
  summarise_at(c("incomepcap","ednum","wid_gini_disp"), mean, na.rm = TRUE) %>% 
  as.data.frame() %>%
  na.omit() %>% 
  dplyr::select(country2,dclass3res_V,incomepcap,ednum,wid_gini_disp) %>% 
  tidyr::pivot_wider(names_from = "dclass3res_V", values_from = c("incomepcap","ednum")) %>% 
  mutate(ratio_class_inc = `incomepcap_Service Class (I+II)`/`incomepcap_Working Class (V+VI+VII)`,
         ratio_class_educ  = `ednum_Service Class (I+II)`/`ednum_Working Class (V+VI+VII)`) %>% 
  as.data.frame()

df_class3_ineq %>% select(wid_gini_disp,ratio_class_inc,ratio_class_educ) %>% tab_corr()

df_class6_ineq <- 
df2 %>%   
  group_by(country2,dclass6res) %>%
  summarise_at(c("incomepcap","ednum","wid_gini_disp"), mean, na.rm = TRUE) %>% 
  as.data.frame() %>%
  na.omit() %>% 
  dplyr::select(country2,dclass6res,incomepcap,ednum,wid_gini_disp) %>% 
  tidyr::pivot_wider(names_from = "dclass6res", values_from = c("incomepcap","ednum")) %>% 
  mutate(ratio_class_inc = `incomepcap_Upper Service class`/`incomepcap_Unskilled working class`,
         ratio_class_educ  = `ednum_Upper Service class`/`ednum_Unskilled working class`) %>% 
  as.data.frame() 
df_class_ineq %>% select(wid_gini_disp,ratio_class_inc,ratio_class_educ) %>% tab_corr()
```

```{r}

income_data

# Example data
# Assuming you have a data frame called 'income_data' with columns: country, social_class, and average_income

# Calculate all pairs of income ratios within each country
income_ratios <- income_data %>%
  group_by(country) %>%
  mutate(income_ratio = combn(average_income, 2, FUN = function(x) x[1] / x[2])) %>%
  ungroup()

# Print the result
print(income_ratios)

>>>>

<!-- We calculated this statistic through a simple earnings regression controlling for age, age squared, and occupation (ISCO 2-digits) in large samples (EU-SILC), or directly for each class in smaller samples (ISSP, EU-EQ), based on the weighted individual data from the IDS; we made these calculations separately for men and women. Based on predicted wage rates, we calculated average hourly wages for each class and took the ratio thereof for different OD class pairs in order to obtain the between-class wage inequality measure. In order to minimize the role of yearly fluctuations (or the 2007-2008 financial crisis), we took the average of yearly ratios available over the first decade of the 21st century. -->

<!-- First, determine the total number of hours worked by multiplying the hours per week by the number of weeks in a year (52). Next, divide this number from the annual salary. For example, if an employee has a salary of $50,000 and works 40 hours per week, the hourly rate is $50,000/2,080 (40 x 52) = $24.04. -->

```{r classineq-data}
summary(df1$incomer)
summary(df1$egp11)
library(DIGCLASS)
frq(df1$WRKHRS)
df1$weekhour <- car::recode(df1$WRKHRS,"c(0,95,96,98,99)=NA")
summary(df1$weekhour); hist(df1$weekhour)
# weekhour * 4 =  monthly worked hours
df1$monthhour <- df1$weekhour*4
countries <- unique(countrycode::countrycode(df1$country,"iso3n","iso3c"))
df1$hhincome <- as.numeric(df1$income/sqrt(df1$hhpop)) # Ingreso per capita del hogar
summary(df1$hhincome)
# Wages for Males
males<- df1%>% 
  filter(female=="Male" & agenum %in% c(18:55)) %>% 
  mutate(country3=countrycode::countrycode(country,"iso3n","iso3c")) %>% 
  select(id,incomer=hhincome,agenum,isco88r,WEIGHT,country3) %>% 
  mutate(isco88_2=factor(isco88_swap(isco88r,from = 4,to = 2)),
         age2=agenum*agenum) %>% na.omit()

# Step 1: Model wages: income regressed age, age squared and occupation
wages_male <- list() 
for (i in countries) {
wages_male[[i]] <-  lm(incomer~agenum+age2+isco88_2,data = males,subset = (country3==i),weights = WEIGHT)
}

# Step 2: predicted wages

predicted_male <- list()
for (i in countries) {
  predicted_male[[i]] <- data.frame(wages=predict.lm(wages_male[[i]]),country=i)
  }
# Step 3: save predicted values
wages_male_df<- bind_rows(predicted_male);dim(wages_male_df)
males$wages_male<- wages_male_df$wages
summary(males$wages_male);hist(males$wages_male)
table(males$wages_male>=0)

# Wages for Females
females<- df1 %>% 
  filter(female=="Female" & agenum %in% c(18:55)) %>%
  mutate(country3=countrycode::countrycode(country,"iso3n","iso3c")) %>%   
  select(id,incomer,agenum,isco88r,WEIGHT,country3) %>% 
  mutate(isco88_2=factor(isco88_swap(isco88r,from = 4,to = 2)),
         age2=agenum*agenum) %>% na.omit() 
# Step 1: Model wages: income regressed on education, age age squared and occupation
wages_female <- list() 
for (i in countries) {
wages_female[[i]] <-  lm(incomer~agenum+age2+isco88_2,data = females,subset = (country3==i),weights = WEIGHT)
}

# Step 2: predicted wages 
predicted_female <- list()
for (i in countries) {
  predicted_female[[i]] <- data.frame(wages=predict(wages_female[[i]]),country=i)
  }
# Step 3: save predicted values
wages_female_df<- bind_rows(predicted_female)
females$wages_female<- wages_female_df$wages
summary(females$wages_female);hist(females$wages_female)
table(females$wages_female>=0)

# Step 4 create the wage, male female dataset to merge to the main data
wages_df<- 
bind_rows(
females %>% select(id,wages=wages_female),
males %>% select(id,wages=wages_male)
)
#Step 5 merge predicted wages to currentdataset
df1 <- left_join(df1,wages_df,"id")

#Step 6: calculate the weekly per hour wage
df1$wageratio <- df1$wages/df1$monthhour
summary(df1$wageratio);hist(df1$wageratio)
table(df1$wageratio>=0)
df1$wageratio[df1$wageratio<0] <- NA # wage ratio per hour lower than 0 is NA
summary(df1$wageratio);hist(df1$wageratio)

# Final: create the dataset for the 6 classes and respective income, education and wage averages

df_hhincome <- 
df1 %>%   
  group_by(country,dclass6res) %>%
  summarise(
    n=n(),
    "incomer"=mean(hhincome,na.rm = TRUE)) %>% 
  as.data.frame() %>% 
  na.omit()

df_education <- 
df1 %>%   
  filter(agenum %in% c(18:64)) %>% 
  group_by(country,dclass6res) %>%
  summarise("ednum"=mean(edyears,na.rm = TRUE)) %>%
  ungroup() %>% 
  # select(-country,-dclass6res) %>%
  as.data.frame() %>% 
  na.omit()

df_wages <- 
df1 %>%   
  group_by(country,dclass6res) %>%
  summarise("wageratio"=mean(wageratio,na.rm = TRUE)) %>%
  ungroup() %>% 
  # select(-country,-dclass6res) %>%
  as.data.frame() %>% 
  na.omit()

table(df_hhincome$country)
table(df_education$country)
table(df_wages$country)

df_class3_ineq<- bind_cols(df_hhincome,df_education %>% select(ednum),df_wages %>% select(wageratio))
df_class3_ineq$country <- countrycode::countrycode(df_class3_ineq$country,"iso3n","iso3c") 
summary(df_class3_ineq)
```

```{r template, eval=FALSE, include=FALSE}
# Step 1: Create the vector of average incomes
average_incomes <- c(upper = 1000, routine = 700, self_employed = 900, working = 500)

# Step 2: Create the matrix of ratios
income_diff_matrix <- outer(average_incomes, average_incomes, FUN = function(x, y) y/x)

# Step 3: Set the diagonal to 0
diag(income_diff_matrix) <- 0

# Step 4: Flatten the matrix, excluding the diagonal
income_diff_vector <- income_diff_matrix[lower.tri(income_diff_matrix, diag = FALSE)] # Extract lower triangular matrix
# income_diff_vector <- c(income_diff_vector, income_diff_matrix[upper.tri(income_diff_matrix, diag = FALSE)]) # Add upper triangular matrix

# Step 5: Calculate the mean of the differences
average_value <- mean(income_diff_vector)

# Display the average value
print(average_value)
```

```{r class-inq-income}
# Step 0: prepare the data
countries<- unique(df_class3_ineq$country)

# Step 1: Create the vector of average incomes
average_incomes <- df_class3_ineq %>% select(country,dclass6res,incomer)

# Step 2: Create the matrix of ratios
matrix_incomer <- list()
for (i in countries) {
  matrix_incomer[[i]] <- outer(subset(average_incomes,country==i)$incomer,subset(average_incomes,country==i)$incomer, FUN = function(x, y) y/x)
}
# Step 3: Set the diagonal to 0
for (i in countries) {
  diag(matrix_incomer[[i]]) <- 0
}
# Step 4: Flatten the matrix, excluding the diagonal
for (i in countries) {
  matrix_incomer[[i]] <- matrix_incomer[[i]][lower.tri(matrix_incomer[[i]], diag = FALSE)] # Extract lower triangular matrix
}
# Step 5: Calculate the mean of the differences
average_incomeratio <- list()
for (i in countries) {
  average_incomeratio[[i]] <- mean(matrix_incomer[[i]])
}

# Display the average value
print(average_incomeratio)
```

```{r class-inq-education}
# Step 0: prepare the data
countries<- unique(df_class3_ineq$country)

# Step 1: Create the vector of average incomes
average_education <- df_class3_ineq %>% select(country,dclass6res,ednum)

# Step 2: Create the matrix of ratios
matrix_education <- list()
for (i in countries) {
  matrix_education[[i]] <- outer(subset(average_education,country==i)$ednum,subset(average_education,country==i)$ednum, FUN = function(x, y) y/x)
}
# Step 3: Set the diagonal to 0
for (i in countries) {
  diag(matrix_education[[i]]) <- 0
}
# Step 4: Flatten the matrix, excluding the diagonal
for (i in countries) {
  matrix_education[[i]] <- matrix_education[[i]][lower.tri(matrix_education[[i]], diag = FALSE)] # Extract lower triangular matrix
}
# Step 5: Calculate the mean of the differences
average_educratio <- list()
for (i in countries) {
  average_educratio[[i]] <- mean(matrix_education[[i]])
}

# Display the average value
print(average_educratio)
```

```{r class-inq-wages}
# Step 0: prepare the data
countries<- unique(df_class3_ineq$country)

# Step 1: Create the vector of average incomes
average_wages <- df_class3_ineq %>% select(country,dclass6res,wageratio)

# Step 2: Create the matrix of ratios
matrix_wages <- list()
for (i in countries) {
  matrix_wages[[i]] <- outer(subset(average_wages,country==i)$wageratio,subset(average_wages,country==i)$wageratio, FUN = function(x, y) y/x)
}
# Step 3: Set the diagonal to 0
for (i in countries) {
  diag(matrix_wages[[i]]) <- 0
}
# Step 4: Flatten the matrix, excluding the diagonal
for (i in countries) {
  matrix_wages[[i]] <- matrix_wages[[i]][lower.tri(matrix_wages[[i]], diag = FALSE)] # Extract lower triangular matrix
}
# Step 5: Calculate the mean of the differences
average_wageratio <- list()
for (i in countries) {
  average_wageratio[[i]] <- mean(matrix_wages[[i]])
}

# Display the average value
print(average_wageratio)
```

```{r}
df_classdiff<- data.frame(t(bind_rows(average_wageratio,average_incomeratio,average_educratio)))
df_classdiff <- 
df_classdiff %>% 
  mutate(country= row.names(df_classdiff),
         ccode= countrycode::countrycode(country, origin = "iso3c",destination = "iso3n")) %>% 
  select(ccode,difclass_wage=X1,difclass_income=X2,difclass_educ=X3)

psych::alpha(x = df_classdiff[,c("difclass_wage","difclass_income","difclass_educ")])
psych::fa(df_classdiff[,c("difclass_wage","difclass_income","difclass_educ")])
model1 <- 'classineq=~difclass_wage+difclass_income+difclass_educ'
fit1<- lavaan::cfa(model = model1,data = df_classdiff)
lavaan::summary(fit1,standardized=T,fit.measures=T)

df_classdiff$difclass <- df_classdiff$difclass_wage + df_classdiff$difclass_income + df_classdiff$difclass_educ
skimr::skim(df_classdiff)
```

```{r merge}
macro_data <- 
  left_join(wiid_2017,solt_2017) %>%
  left_join(pen_2017) %>% 
  left_join(otero) %>%
  left_join(ocde) %>% 
  left_join(wid_df1) %>% 
  left_join(ilo_df) %>% 
  left_join(df_classdiff)

correlation::correlation(macro_data)
save(macro_data,file = here::here("input/data/proc/study1_macro_data.RData"))

@@
